---
title: "Black Politicians Training Aid: Evaluated Code"
format:
  html:
    toc: true
    code-links:
      - text: Training Aid
        href: black_politicians_lab.html
      - text: Full Code
        href: black_politicians_lab_full_code.html
      - text: QA Log
        href: black_politicians_lab_qa_log.html
      - text: Evaluated Script
        href: black_politicians_lab_evaluated_code.R
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

This page runs the evaluated workflow script and shows processed data objects created during execution.

```{r}
#| include: false
source("black_politicians_lab_evaluated_code.R", local = knitr::knit_global())
```

## Evaluated Script

```{r}
#| results: asis
cat("```r\n")
cat(readLines("black_politicians_lab_evaluated_code.R"), sep = "\n")
cat("\n```")
```

## Processed Data Preview

```{r}
dat_preview
```

## Raw Comparison Summary

```{r}
raw_summary
```

## Matched Sample Counts

```{r}
exact_counts
cem_counts
```

## Entropy Balancing Weights Summary

```{r}
ebal_weight_summary
```

## Effect Estimates By Design

```{r}
estimate_comparison
```

## Book-Replication Interaction Model (CEM Weights)

```{r}
final_fit_key_terms
```
