---
title: "HISP IE in Practice: Evaluated Code"
format:
  html:
    toc: true
    code-links:
      - text: Lab
        href: hisp_ie_practice_lab.html
      - text: Full Code
        href: hisp_ie_practice_lab_full_code.html
      - text: QA Log
        href: hisp_ie_practice_lab_qa_log.html
      - text: Evaluated Script
        href: hisp_ie_practice_lab_evaluated_code.R
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

This page runs the evaluated HISP workflow script and shows processed outputs for each branch of the translated Stata replication.

```{r}
#| include: false
source("hisp_ie_practice_lab_evaluated_code.R", local = knitr::knit_global())
```

## Evaluated Script

```{r}
#| results: asis
cat("```r\n")
cat(readLines("hisp_ie_practice_lab_evaluated_code.R"), sep = "\n")
cat("\n```")
```

## CSV Export Manifest

```{r}
csv_export_manifest
```

## Data Overview

```{r}
data_overview
panel_completeness
data_preview
```

## Effect Summary Across The Do-File Sequence

```{r}
effect_pathway_summary
```

## Mean-Comparison Checks

```{r}
method_1_ttest
method_2_ttest
method_3_outcome_ttests
method_6_ttests
```

## Randomized-Assignment Baseline Balance

```{r}
method_3_baseline_control_ttests
```

## Instrumental-Variables Branch

```{r}
method_4_ttests
method_4_first_stage
method_4_iv_regressions
```

## Regression-Discontinuity Branch

```{r}
method_5_regressions
```

```{r}
rdd_density_plot
rdd_participation_plot
rdd_followup_plot
```

## Matching Branch

```{r}
matching_wide_summary
matching_restricted_support
matching_restricted_regression[matching_restricted_regression$term == "enrolled", ]
matching_restricted_bootstrap
matching_full_support
matching_full_regression[matching_full_regression$term == "enrolled", ]
matching_full_bootstrap
matched_did_manual_summary
matched_did_regression[matched_did_regression$term == "enrolled", ]
```

```{r}
matching_restricted_ps_plot
matching_full_ps_plot
```

## Power Calculations

```{r}
power_parameter_summary
power_health_no_cluster_90
power_health_no_cluster_80
power_hospital_no_cluster_80
power_health_clustered_100
power_health_clustered_varying_clusters
```
