---
title: "LaLonde Main Report Lab: Evaluated Code"
format:
  html:
    toc: true
    code-links:
      - text: Lab
        href: lalonde_main_report_lab.html
      - text: Main Report
        href: ../articles/matching_methods_report.html
      - text: Full Code
        href: ../articles/matching_methods_report_full_code.html
      - text: QA Log
        href: ../articles/matching_methods_report_qa_log.html
      - text: Evaluated Script
        href: lalonde_main_report_lab_evaluated_code.R
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

This page runs the evaluated LaLonde main-lab script and shows the key processed outputs used in the guided worksheet.

```{r}
#| include: false
source("lalonde_main_report_lab_evaluated_code.R", local = knitr::knit_global())
```

## Evaluated Script

```{r}
#| results: asis
cat("```r\n")
cat(readLines("lalonde_main_report_lab_evaluated_code.R"), sep = "\n")
cat("\n```")
```

## Processed Data Preview

```{r}
dat |>
  dplyr::select(treat, outcome, dplyr::all_of(design_covariates)) |>
  head(10)
```

## Raw Comparison Summary

```{r}
pre_match_summary
```

## Baseline Balance Benchmark

```{r}
unadjusted_balance
```

## Exact Matching Outputs

```{r}
exact_retention
exact_effect
```

## CEM Outputs

```{r}
cem_comparison
cem_effect
```

## Entropy-Balancing Outputs

```{r}
ebal_tuning_comparison
ebal_weight_diagnostics
ebal_effect
```

## Side-By-Side Comparison Tables

```{r}
comparison_balance
comparison_information
treatment_effect_summary
```
