---
title: "Lab: Titanic Teaching Example"
format:
  html:
    code-links:
      - text: Full Code
        href: titanic_teaching_lab_full_code.html
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

```{r}
#| include: false
source("titanic_teaching_lab_full_code.R", local = knitr::knit_global())
```

## How To Use This Page

This lab turns the Titanic scaffold into a compact executed example for exact matching and propensity-score subclassification.

::: {.callout-note}
This is a teaching example, not a literal policy-evaluation design. The value of `causaldata::titanic` is that the treatment groups, covariates, and support cells are easy to see on one page.
:::

## Design Choice

For a slide-friendly matching example, keep the setup deliberately simple:

- Treatment: `sex`, coded here as female versus male
- Outcome: `survived`
- Pre-treatment covariates: ticket `class` and `age`
- Teaching estimand: `ATT` for the female group

This choice keeps the design transparent. There are only two adjustment variables, both are discrete, and the implied exact-match cells are easy to inspect directly before using any propensity-score machinery.

## Dataset At A Glance

In this local installation, `causaldata::titanic` contains `r format(nrow(dat), big.mark = ",")` passenger records with four variables: `class`, `age`, `sex`, and `survived`.

The table below shows the raw comparison before any design adjustment.

```{r}
knitr::kable(
  raw_summary_display,
  align = c("l", "r", "r", "r", "r", "r"),
  caption = "Raw comparison: women and men differ on both survival and covariate composition."
)
```

Women survive at a much higher rate than men in the raw data, but women are also less concentrated in the crew and somewhat more concentrated in the higher-status passenger classes. That makes this a useful classroom example for separating the raw contrast from the adjusted one.

## Support Cells Before Matching

There are `r total_cells_n` possible age-by-class cells. `r observed_cells_n` are observed in the data, and all `r supported_cells_n` observed cells contain both women and men. The only empty cell is `r empty_cells_text`.

```{r}
support_plot
```

```{r}
knitr::kable(
  cell_support_display,
  align = c("l", "l", "r", "r", "r", "r", "r", "l"),
  caption = "Exact-match cells: counts and within-cell survival rates."
)
```

This is why Titanic works well in teaching. You can inspect the exact cells first and see that exact matching is feasible before fitting a propensity model.

## Exact Matching

Exact matching on `class` and `age` is almost the ideal first demonstration here:

- every observed female passenger has at least one male comparison unit in the same cell
- the matched sample achieves exact balance on the design covariates
- the estimand remains easy to explain because the cells are visible

```{r}
knitr::kable(
  balance_compare_display,
  align = c("l", "r", "r", "r"),
  caption = "Absolute standardized mean differences before adjustment and after each design."
)
```

The exact-matching column is all zeros because the design conditions directly on the full set of observed covariates.

## Propensity-Score Subclassification

Subclassification is useful here as a bridge to higher-dimensional problems. With only two discrete covariates it does not buy much over exact matching, but it shows how a propensity-score design can recover almost the same adjusted comparison while relaxing the requirement that each design cell be used literally.

```{r}
knitr::kable(
  subclass_counts_display,
  align = c("l", "r", "r"),
  caption = "Subclass composition after propensity-score subclassification."
)
```

All four subclasses retain both women and men, and the adjusted balance remains well below the `0.1` standardized-difference benchmark on every covariate.

## Comparison Of Estimates

```{r}
knitr::kable(
  estimate_table_display,
  align = c("l", "r", "r", "r", "r"),
  caption = "Raw and adjusted survival differences for the female group."
)
```

The key teaching result is that adjustment changes the estimate only modestly:

- raw difference: `r sprintf('%.1f', raw_gap_pp)` percentage points
- exact matching: `r sprintf('%.1f', exact_gap_pp)` percentage points
- subclassification: `r sprintf('%.1f', subclass_gap_pp)` percentage points

That is a useful contrast with higher-dimensional matching examples. Here, the adjustment task is so transparent that the class-by-age structure already explains most of what the design is doing.

## Slide-Friendly Summary

1. Start with the cell structure. Titanic is useful because you can inspect the exact comparison cells directly.
2. Exact matching is feasible here because the covariates are low-dimensional and discrete.
3. Subclassification reaches nearly the same adjusted estimate, which makes it a clean bridge to settings where exact matching becomes too restrictive.
4. This is still a pedagogic example rather than a literal causal treatment effect, because `sex` is not a manipulable intervention and important determinants of survival are not observed.
